### 2024-09-21 paper/paper.qmd
Initial Commit
template from IC3

---
title: "Marriage Licence Analysis"
subtitle: "STA304 In-class 3"
author: 
  - Amy Jin
thanks: "Code and data are available at: https://github.com/aj3616/STA304IC3"
date: today
date-format: long
abstract: "Data analysis for marrige licence statistics. Graphs and text produced in quarto"
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

# Introduction

This is a analysis about Marriage Licence in different district in Canada, We use @talia, @opendatatorotno, and @rohan.

The remainder of this paper is structured as follows. @sec-mydatasection

# Data {#sec-mydatasection}

Some of our data is of penguins (@fig-marriage), from @palmerpenguins.

```{r}
#| label: fig-marriage
#| fig-cap: Bills of penguins
#| echo: false

data |> 
  ggplot(aes(x = date, y = marriage_licenses)) +
  geom_point()

```

# Discussion

## First discussion point {#sec-first-point}
This is the scatterplot generated by the random sample we selected, this shows an ocsilating trend for marriage licneces
$[0, 500]_{marraigelicence} \land [2010, 2020]_{date}$ and during 2020 and 2025, there is an increasing number of marraige licences and the scatterplots did not clustered in $[0, 500]_{marraigelicence}$.

## Second discussion point
It is safe to assume that the total number of marriages did not vary much, but the average number of licence did increase.

## Third discussion point
It is arguable that less people are getting married in the contemporary time.

## Weaknesses and next steps
We need to perform more statistics analysis to make a more valid conclusion, some next steps can be Times Series analysis, regression estimates and visual representations.

\newpage

\appendix

# Appendix {-}

# Additional data details

\newpage

# References