### 2024-09-24 paper/paper.qmd
first draft of paper & readme

---
title: "Non-Regulated Residential Lead Testing Program in Toronto"
subtitle: "Lead Contamination in Water"
author: 
  - Amy Jin
thanks: "Code and data are available at: https://github.com/aj3616/Lead-Testing-Program-in-Toronto"
date: today
date-format: long
abstract: "The City of Toronto offers a free, non-regulated Residential Lead Testing Program, enabling residents to collect and submit water samples for lead testing. Quality of drinking water is an important public health issue. This paper examines how time and district affects the lead concentration with respect to safe lead concentration guidelines. It also highlights ongoing steps being taken to mitigate lead contamination in the water supply."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
lead = data$`Lead Amount (ppm)`

```

# Introduction

Lead contamination in drinking water poses significant public health
risks, particularly for vulnerable populations such as children and
pregnant women. Even low levels of lead exposure can have serious health
consequences, including developmental delays and cognitive impairments.
In response to growing concerns about lead exposure, the City of Toronto
has implemented a non-regulated Residential Lead Testing Program. This
initiative allows residents to access free water testing services,
providing them with the means to assess the lead content in their tap
water.

The testing process is straightforward: residents can obtain a testing
kit from any of six Toronto Public Health locations, collect a water
sample from their home, and return it for analysis. Toronto Water
processes the samples and provides results directly to the residents.
Although the City cannot control how the samples are collected, this
service empowers residents to take action in ensuring the safety of
their drinking water. This paper will explore the benefits of the
program, its limitations, and the broader efforts undertaken by the City
of Toronto to address lead contamination in its water supply system.

We use @talia, @opendatatorotno, and @rohan.

The remainder of this paper is structured as follows. @sec-mydatasection

# Data {#sec-mydatasection}

```{r}
#| label: table1summary
#| fig-cap: "Table 1: Summary of the lead amount in ppm"
#| echo: false
summary(data$`Lead Amount (ppm)`)
```

```{r}
#| label: figure1boxplot
#| fig-cap: "Figure 1: summary of the lead amount in ppm"
#| echo: false
boxplot(data$`Lead Amount (ppm)`)
abline(h = 0.010011423, col = "red")
```

```{r}
#| echo: false
threshold <- 0.010011423
below = data$`Lead Amount (ppm)` < threshold
sum(below)/10215

```

```{r}
#| label: figure2ggplot
#| fig-cap: "Figure 2: Lead Levels changed over time for times seriese analysis"
#| echo: false
plot(data$`Sample Date`, data$`Lead Amount (ppm)`, type='l', main="Lead Levels Over Time", xlab="Date", ylab="Lead Amount (ppm)")

```

```{r}
#| label: figure3ACF
#| fig-cap: "Figure 3: Autocorrelation function for lead concentrations"
#| echo: false
# Autocorrelation function (ACF)
acf(data$`Lead Amount (ppm)`, na.action = na.omit, main="Autocorrelation of Lead Levels")
```

```{r}
#| label: figure4PACF
#| fig-cap: "Figure 4: Partial Autocorrelation function for lead concentrations"
#| echo: false
# Partial Autocorrelation function (PACF)
pacf(data$`Lead Amount (ppm)`, na.action = na.omit, main="Partial Autocorrelation of Lead Levels")

```

```{r}
#| label: figure5
#| fig-cap: "Figure 5: Outliers categorized by district"
#| echo: false
boxplot(`Lead Amount (ppm)` ~ PartialPostalCode, data=data, main="Lead Levels by Postal Code", xlab="Partial Postal Code", ylab="Lead Amount (ppm)")

```

```{r}
#| label: figure6
#| fig-cap: "Figure 6: Mean lead concentration categorized by first two characters in partial postal code"
#| echo: false
# Load necessary library
library(ggplot2)

# Remove rows with LeadAmountPPM > 4
filtered_data <- subset(data, `Lead Amount (ppm)` <= 4)

# Categorize by the first two characters of PartialPostalCode
filtered_data$District <- substr(filtered_data$PartialPostalCode, 1, 2)

# Step 3: Calculate the mean of LeadAmountPPM by District
mean_lead_by_district <- aggregate(`Lead Amount (ppm)` ~ District, data = filtered_data, FUN = mean)
print(mean_lead_by_district)
# Step 4: Construct a bar graph using ggplot2
ggplot(mean_lead_by_district, aes(x = District, y = `Lead Amount (ppm)`)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Mean Lead Concentration by District",
       x = "District (First Two Characters of Postal Code)",
       y = "Mean Lead Amount (ppm)") +
  theme_minimal()

```

```{r}
threshold <- 0.010
high_lead_samples <- data[data$`Lead Amount (ppm)` > threshold, ]

# Count of samples above the threshold
nrow(high_lead_samples)

# Percentage of samples above threshold
(nrow(high_lead_samples) / nrow(data)) * 100

```

# Discussion

## Lead Concentration influencing Health

-   WHO and Canadian federal guidance on lead concentration in water

Lead amount The main variable we are examining is the Lead Amount in
ppm. The distribution of the data is highly right skewed with most
of its data being very small, as shown in figure 1(@figure1boxplot),
most data are almost 0 with some outliers which are mostly less than
2 ppm and 4 of them ranging from 3 to 8.44 which is the maximum
value as shown in the summary table(@table1summary). We can also
verify our observation since the summary(@table1summary) shows that
75% of the data is below 0.001480. Government of Canada had
specified that the MAC of drinking water is 0.010mg/L(Government
Canada), which converts to 0.010011423
part/million(unitconverters.net), this is shown by the red line in
the figure 1(@figure1boxplot).

-   percentage of data above this threshold

From calculations, there are 97.26% of the data collected indicatin that the lead amount in the water is safe, and it is safe to assume that 75% of the areas in toronto is
safe of lead in water.

-   plot of the data above this threshold

## How districts affect Lead Concentration

-   graphs of postal code vs lead concentrations

Postal Code Figure 5(@figure5) is a scatterplot of each specific
partial postal code(first 3 code), we can see that again, most data
points are around 0, but there are a few outliers with very high
lead concentration. Figure 5 clarily indicated that which district
are those outliers in. We can see that the "M4" has two datapoints
aroun 8 ppm, as well as some between 0 and 2ppm. "M6" also have many
data between 0 and 2ppm. It is safe to assume that those two district 
might have a higher overall lead concentration level.

-   extract outliers above the threshold

We will verify the assumption with calculating the mean lead
concentration when categorized the partial postal codes by only the
first two numbers. Since outliers might influence the mean, especially
when in our dataset we are dealing with median 0.000473 and outliers as
high as 8ppm, we can omit the outliers above 4ppm and perform the data
analysis.

-   mean value of lead concentration for different districts

Figure 6(@figure6) is the bargraph comparing the mean values
of lead concentration without outliers above 4ppm and cetegorized by
only the first two characters of the partial postal code.
We can see that the top 3 district of the highest lead concentration in
water is M9, M6 and M5 respectively. This shows that without the data
above 4ppm,

-   toronto map

## How months and year affect Lead Concentration

-   times series analysis for autocorrelation

Time and Lead In this dataset, we have variables time and lead amount, therefore, we can perform times series analysis(STA457). As shown in figure 2(@figure2ggplot), we can see that except for some sharp increase in 2014, 2019, 2021, 2023, the overall trend is stays
around 0 and remains positive.

From the ACF and PACF figures, we can conclude that the previous values
of lead concentrations does not impact future values. For ACF, when lag
is 0, ACF is always 1 but for all other lags in ACF(@figure3ACF) and
PACF(@figure4PACF) they are all within the dotted blue lines which is
the evidence that there are no significant autocorrelation nor
autoregressive realtionship present in the data.

This observation is reasonable because the samples are collected in
different district in Toronto, and it is safe to assume that one
incident of a very high lead concentration might again only be outliers
instead of some factors that will effect the overall water quality in
Toronto.

-   graphs of months vs mean lead concentration

-   graphs of years vs mean lead concentration

-   extract outliers above the threshold

## Weaknesses and next steps

-   the data is from many different water source, they are not
    necessarily direct drinking water since how the water sample
    collected was not controled. However, the guides are for drinking
    water.

-   importance of pre-treatment such as filter

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

\newpage

# References